---
title: "Untitled"
output:
  word_document: default
  html_document: default
  pdf_document: default
---


#  Gene regulatory network

```{r pressure, echo=TRUE}
knitr::opts_chunk$set(echo = TRUE)
library(pcalg)

data <- read.csv("./BRCA_RNASeqv2_top50.csv")
data$class <- ifelse(data$class =="C", 1, 0)
# data$class[data$class=="C"] <- 1
# data$class[data$class=="N"] <- 0
data$class <- as.numeric(data$class)
genes <- subset(data, select=-(class))

n <- nrow (genes)
V <- colnames(genes) 
pc.fit <- pc(suffStat = list(C = cor(genes), n = n),
             indepTest = gaussCItest, alpha=0.01, labels = V)
if (require(Rgraphviz)) {
  ## show estimated graph
  # par(mfrow=c(1,2))
  plot(pc.fit
       , main = "Graph 1 - Causal Structure")
}
```


# How the PC algorithm works
#  10 genes that have strong causal effects on EBF1
```{r message=FALSE, warning=FALSE}
pcS <- pcSelect(genes[c("EBF1")], genes[, !names(genes) %in% c("EBF1")] , alpha=0.05)
# pcS
pcs_frame <- data.frame(pcS,stringsAsFactors = FALSE)
pcs_frame$gene <- rownames(pcs_frame)
rownames(pcs_frame) <- NULL
newdata <- pcs_frame[order(-pcs_frame$zMin),]
newdata[0:10,]
# selected_genes <- sort(abs(pcS$zMin),decreasing = TRUE)[0:10]
# selected_genes
```
# Genes in the Markov blanket
```{r message=FALSE, warning=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library(bnlearn)
MB.Z=learn.mb(genes, "ABCA9", method="iamb", alpha=0.01)
MB.Z 
```
# How the Markov blanket algorithm works


# How PC-simple works


# Parent and children set of the class variable by PC-simple
```{r message=FALSE, warning=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library("caret")
library(e1071)
library("klaR")

mean_all <- mean(as.matrix(genes))
genes <- as.data.frame(ifelse(genes > mean_all, 1, 0))
data1 <- data
data1[, !names(data1) %in% c("class")]  <- as.data.frame(ifelse(data[,!names(data) %in% c("class")]  > mean_all, 1, 0))


pcS4 <- pcSelect(data[c("class")], genes , alpha=0.05)

# selected_genes <- sort(abs(pcS4$zMin),decreasing = TRUE)[0:10]
pcs4_frame <- data.frame(pcS4,stringsAsFactors = FALSE)
pcs4_frame$gene <- rownames(pcs4_frame)
rownames(pcs4_frame) <- NULL
newdata4 <- pcs4_frame[order(-pcs4_frame$zMin),]
pcs_genes <- newdata4[newdata4$G==TRUE,c("gene")]
pcs_genes
```

# Naïve Bayes classification with all features
```{r message=FALSE, warning=FALSE}
# knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
# genes[sapply(genes, is.double)] <- lapply(genes[sapply(genes, is.double)], as.factor)
# genes
data$class <- as.factor(data$class)
train_index <- sample(1:nrow(genes), 0.8 * nrow(genes))
test_index <- setdiff(1:nrow(genes), train_index)
train <- genes[train_index,]
test <- genes[test_index,]
# x <- train[, !names(train) %in% c("class")] 
x <- train
y <- data[train_index,]$class

model = train(x,y,'nb',trControl=trainControl(method='cv',number=10))

x <- test
y <- data[test_index,]$class
confusionMatrix <- prop.table(table(predict(model$finalModel,x)$class,y))
confusionMatrix
           
```
## performance metrics
```{r message=FALSE, warning=FALSE}

accuracy <- (confusionMatrix[1,1]+confusionMatrix[2,2])/
  (confusionMatrix[1,1]+confusionMatrix[1,2]+
     confusionMatrix[2,1]+confusionMatrix[2,2])
recall <-  confusionMatrix[1,1]/
  (confusionMatrix[1,1]+confusionMatrix[1,2])
Precision <-  confusionMatrix[1,1]/
  (confusionMatrix[1,1]+confusionMatrix[2,1])
F1 <- 2*(Precision*recall)/(Precision+recall)
accuracy
recall        
Precision
F1
```
# Naïve Bayes classification with features obtain by PC-simple
```{r message=FALSE, warning=FALSE}

#with selective genes################################

genes_select <- genes[,c(pcs_genes)]

train_index <- sample(1:nrow(genes_select), 0.8 * nrow(genes_select))
test_index <- setdiff(1:nrow(genes_select), train_index)
train <- genes_select[train_index,]
test <- genes_select[test_index,]
x <- train
y <- data[train_index,]$class

model = train(x,y,'nb',trControl=trainControl(method='cv',number=10))

x <- test
y <- data[test_index,]$class
confusionMatrix_select <- prop.table(table(predict(model$finalModel,x)$class,y))
confusionMatrix_select

```
## performance metrics
```{r message=FALSE, warning=FALSE}

accuracy <- (confusionMatrix_select[1,1]+confusionMatrix_select[2,2])/
  (confusionMatrix_select[1,1]+confusionMatrix_select[1,2]+
     confusionMatrix_select[2,1]+confusionMatrix_select[2,2])
recall <-  confusionMatrix_select[1,1]/
  (confusionMatrix_select[1,1]+confusionMatrix_select[1,2])
Precision <-  confusionMatrix_select[1,1]/
  (confusionMatrix_select[1,1]+confusionMatrix_select[2,1])
F1 <- 2*(Precision*recall)/(Precision+recall)
accuracy
recall
Precision
F1
```





# conditional probability tables
```{r message=FALSE, warning=FALSE}
library(dplyr)

library(gRain)

BTNL9_ <- table(data1[,c("BTNL9")])
CD300LG_BTNL9 <- data1[,c("CD300LG", "BTNL9")] %>% group_by_all %>% count
class_CD300LG <- data1[,c("class","CD300LG")] %>% group_by_all %>% count
IGSF10_class <- data1[,c("IGSF10","class")] %>% group_by_all %>% count
ABCA9_IGSF10_BTNL9 <- data1[,c("ABCA9", "IGSF10", "BTNL9")] %>% group_by_all %>% count

lvl <- levels(as.factor(data1$BTNL9))

BTNL9 <- cptable(~BTNL9, values=as.vector(BTNL9_),levels=lvl)
CD300LG <- cptable(~CD300LG|BTNL9, values=CD300LG_BTNL9$n,levels=lvl)
class_ <- cptable(~class_|CD300LG, values=class_CD300LG$n, levels=lvl)
IGSF10 <- cptable(~IGSF10|class_, values=IGSF10_class$n, levels=lvl)
ABCA9 <- cptable(~ABCA9|IGSF10:BTNL9,values=ABCA9_IGSF10_BTNL9$n,levels=lvl)

plist <- compileCPT(list(BTNL9,CD300LG,class_,IGSF10,ABCA9))
# plist
net1=grain(plist)
# plot(net1) # undirected network
plot(net1$dag) # directed network

```

# Probability of the four genes with high expression levels
```{r message=FALSE, warning=FALSE}
querygrain(net1, nodes=c("BTNL9","CD300LG","IGSF10","ABCA9"), type="marginal")
# sum(data1$BTNL9)/length(data1$BTNL9)

```

# probability of having cancer when the expression level of CD300LG is high and BTNL9 is low
```{r message=FALSE, warning=FALSE}
#P(class_=yes|CD300LG=yes, BTNL9=NO)
querygrain(net1, nodes=c("class_","CD300LG","BTNL9"), type="conditional")

```
```{r message=FALSE, warning=FALSE}
#P(class_=yes|CD300LG=yes, BTNL9=NO)
querygrain(net1, nodes=c("class_","CD300LG","BTNL9"), type="conditional")

```
# Prove the result

```{r message=FALSE, warning=FALSE}
#P(class_=yes|CD300LG=yes, BTNL9=NO)
querygrain(net1, nodes=c("class_","CD300LG"), type="conditional")

```

```{r message=FALSE, warning=FALSE}
#P(class_=yes|CD300LG=yes, BTNL9=NO)
querygrain(net1, nodes=c("CD300LG","BTNL9"), type="conditional")

```
```{r message=FALSE, warning=FALSE}
#P(class_=yes|CD300LG=yes, BTNL9=NO)
querygrain(net1, nodes=c("class_","CD300LG","ABCA9"), type="conditional")

```

# Why the “class” conditionally independent of ABCA9





